Functional Pearl: Short and Mechanized Logical Relation for Dependent Type Theories

### 2024-12-18 paper/paper.tex
Remove some handled comments

Our definitional equality is untyped: the judgment does not require the two
terms to type check and have the same type. The use of an untyped relation for
conversion is similar to Barendregt's Pure Type
Systems~\cite{barendregt1991introduction} and differs
from MLTT~\citep{Martin-Lof-1973}, where definitional
equality takes the form $\Gamma \vdash a \equiv b : A$.
% from which one can
% usually derive $[[G |- a : A]]$ and $[[G |- b : A]]$ after proving subject
% reduction.
By working with an untyped judgment, we can establish its
properties independently of the type system and the logical relation, using
well-established syntactic approaches.
%
\citet{siles2012pure} show the equivalence of Barendregt's Pure Type
System, which employs untyped equality, and
its variant that uses typed equality. This assures us that we do
not lose generality working with a system with untyped
conversion.
\jc{I'm not too fond of this last sentence because we \emph{do} lose
the ability to add typed $\eta$ equivalences.}
\yl{removed the eta equivalence sentence because it might give the
  wrong message that we extended Siles' work with eta:
 Furthermore, we discuss how this definition can be extended with
 $\eta$-equivalence of functions in Section~\ref{XXX} }
We compare this
definition with type-directed approaches to equality in
Section~\ref{sec:discuss}.

@@

\scw{Cassia did not follow our discussion of dependent pattern matching. We need to expand}
The elimination form for booleans in \rref{T-If} demonstrates dependent pattern
matching: the type of the expression \emph{depends} on the term being matched.
This result type involves a \emph{motive} $[[A]]$ abstracted over a boolean.
For the overall expression, $[[x]]$ is filled in by the match target $[[a]]$,
while in the $[[true]]$ (resp. $[[false]]$) branch, $[[x]]$ is filled in by
$[[true]]$ (resp. $[[false]]$). Informally, the type of each branch is more
precise than those in a nondependent match expression, since it knows which
branch it is in. An alternative perspective is that to prove $[[A {a/x}]]$, it
suffices to show that it holds for the two canonical cases that $[[a]]$ is
$[[true]]$ or is $[[false]]$.

@@

\scw{Would it make sense to define the notation $a \in [\![A]\!]^i$ when
there exists some $S$ such that $[[InterpR i A S]]$ and $[[a in S]]$ ?}
\jc{I like this notation too, $[[a in InterpR i A]]$}

@@

\scw{This lemma is only used to prove \texttt{wn\_antirenaming}, which is then
used more generally. Maybe we should replace it with  that?}
\yl{Anti-renaming is mentioned once in Section 6 when we talk about how confluence with $\eta$
depends on anti-renaming for Par. I don't think that discussion in
Section 6 is very important so I'm fine with replacing it}

>>>>

Our definitional equality is untyped: the judgment does not require the two
terms to type check and have the same type. The use of an untyped relation for
conversion is similar to Barendregt's Pure Type
Systems~\cite{barendregt1991introduction} and differs
from MLTT~\citep{Martin-Lof-1973}, where definitional
equality takes the form $\Gamma \vdash a \equiv b : A$.
% from which one can
% usually derive $[[G |- a : A]]$ and $[[G |- b : A]]$ after proving subject
% reduction.
By working with an untyped judgment, we can establish its
properties independently of the type system and the logical relation, using
well-established syntactic approaches.
%
\citet{siles2012pure} show the equivalence of Barendregt's Pure Type
System, which employs untyped equality, and
its variant that uses typed equality. This assures us that we do
not lose generality working with a system with untyped
conversion.
We compare this
definition with type-directed approaches to equality in
Section~\ref{sec:discuss}.

@@

The elimination form for booleans in \rref{T-If} demonstrates dependent pattern
matching: the type of the expression \emph{depends} on the term being matched.
This result type involves a \emph{motive} $[[A]]$ abstracted over a boolean.
For the overall expression, $[[x]]$ is filled in by the match target $[[a]]$,
while in the $[[true]]$ (resp. $[[false]]$) branch, $[[x]]$ is filled in by
$[[true]]$ (resp. $[[false]]$). Informally, the type of each branch is more
precise than those in a nondependent match expression, since it knows which
branch it is in. An alternative perspective is that to prove $[[A {a/x}]]$, it
suffices to show that it holds for the two canonical cases that $[[a]]$ is
$[[true]]$ or is $[[false]]$.

@@

@@
